---
title: "Valdivia et al 2021"
author: "Tania Valdivia Carrillo"
date: "04/11/2022"
output:
  pdf_document: default
  html_document: default
---

```{r}
# Load the required packages
library(ade4)
library(adespatial)
library(vegan)
library(gclus)
library(cluster)
library(FD)
```

DATA
```{r}
edna_reads<-read.csv("edna.csv",row.names = 1)
```

UVC
```{r}
data_UVC<-read.csv("uvc.csv",row.names = 1)
```

SPATIAL
```{r}
spa<-read.csv("spa.csv", row.names = 1)
```

ENVIRONMENTAL
```{r}
env<-read.csv("env.csv", row.names = 1)
```

-----------------------------------------
Abundance distribution
-----------------------------------------
count abundances eDNA
```{r}
#Count cases for each abundance class
ab<-table(unlist(edna_reads))
hist(log10(ab), xlab="Abundance class")
```

```{r}
#Barplot of the distribution , all species
barplot(ab, las=1, log="y", xlab="Read Abundance", ylab="Frequency", col="gray")
```


count abundances UVC
```{r}
#Count cases for each abundance class
ab_UVC<-table(unlist(data_UVC))
hist(log10(ab_UVC), xlab="Abundance class")
```

```{r}
#Barplot of the distribution , all species confoundes
barplot(ab_UVC, las=1, log="y", xlab="Ind Abundance", ylab="Frequency", col="gray")
```

----------------
Read abundance / specie, Ind abundance/specie
----------------
```{r}
spe_presence<-apply(edna_reads,2,sum)
spe_presence<-sort(spe_presence)
plot(spe_presence, pch=16, col="black", cex=0.5, ylab="Read abundace", xlab="OTU", log="y")
```
```{r}
spe_presence_UVC<-apply(data_UVC,2,sum)
spe_presence_UVC<-sort(spe_presence_UVC)
plot(spe_presence_UVC, pch=16, col="black", cex=0.5, ylab="Ind abundace", xlab="Species", log="y")
```
-------------
DIVERSITY INDICES
-----------------

```{r}
#Species richness
S.eDNA<-rowSums(edna_reads>0)
#Shannon entropy
H<-diversity(edna_reads)
#Shannon diversity number
N1<-exp(H)
#Eveness
J <- H / log(S.eDNA)
(diversity_index_edna<-data.frame(S.eDNA,H,N1,J))
```

```{r}
S.UVC<-rowSums(data_UVC>0)
#Shannon entropy
H.UVC<-diversity(data_UVC)
#Shannon diversity number
N1.UVC<-exp(H.UVC)
#EVENESS
J.UVC <- H.UVC / log(S.UVC)
```


```{r}
(diversity_index_edna<-data.frame(S.UVC,H,N1,J))
```

```{r}
(diversity_index_uvc<-data.frame(S.UVC,H.UVC,N1.UVC,J.UVC))
```
Plot

```{r}
sit.pres<-apply(edna_reads>0,1, sum)
(sit.pres.sorted<-sort(sit.pres))

plot(sit.pres, type="h", ylim=c(10,60),las=1, col="royalblue3",cex.main=1, main="Observed OTUs eDNA", cex.lab=1, cex.axis=1, ylab="Sobs",xlab="Sites", family="Helvetica")
text(sit.pres, row.names(edna_reads), cex=0.55, pos=3,col="black", family="Helvetica")
```

```{r}
sit.pres.UVC<-apply(data_UVC>0,1, sum)
(sit.pres.sorted.UVC<-sort(sit.pres.UVC))
plot(sit.pres.UVC, type="h", ylim=c(10,60),las=1, col="royalblue3",cex.main=1, main="Observed species UVC", cex.lab=1, cex.axis=1, ylab="Sobs",xlab="Sites", family="Helvetica")
text(sit.pres.UVC, row.names(data_UVC), cex=0.55, pos=3,col="black", family="Helvetica")
```

```{r}
## Compare sites: species richness
# Compute the number of species at each site
# To sum by rows, the second argument of apply(), MARGIN, is
# set to 1
sit.pres <- apply(edna_reads > 0, 1, sum)
# Sort the results in increasing order
sort(sit.pres)
par(mfrow = c(1, 2))
# Plot species richness vs. position of the sites along the river
plot(sit.pres,type = "s",
las = 1,
col = "gray",
main = "Sobs/site eDNA",
xlab = "Sites",
ylab = "Sobs"
)
text(sit.pres, row.names(edna_reads), cex = .6, col = "black")
# Use geographic coordinates to plot a bubble map
plot(spa,
asp = 1,
main = "Sobs Latitudinal gradient eDNA",
pch = 21,
col = "white",
bg = "red",
cex =   (sit.pres / min(sit.pres))^2,
xlab = "x coordinate",
ylab = "y coordinate"
)
lines(spa, col = "light blue")
```

```{r}
## Compare sites: species richness
# Compute the number of species at each site
# To sum by rows, the second argument of apply(), MARGIN, is
# set to 1
sit.pres.UVC <- apply(data_UVC > 0, 1, sum)
# Sort the results in increasing order
sort(sit.pres.UVC)
par(mfrow = c(1, 2))
# Plot species richness vs. position of the sites along the river
plot(sit.pres.UVC,type = "s",
las = 1,
col = "gray",
main = "Sobs/site UVC",
xlab = "Sites",
ylab = "Sobs"
)
text(sit.pres.UVC, row.names(edna_reads), cex = .6, col = "black")
# Use geographic coordinates to plot a bubble map
plot(spa,
asp = 1,
main = "Sobs Latitudinal gradient UVC",
pch = 21,
col = "white",
bg = "red",
cex =   (sit.pres.UVC / min(sit.pres.UVC)),
xlab = "x coordinate",
ylab = "y coordinate"
)
lines(spa, col = "light blue")
```

```{r}
#build a expected curve (randomization for boxplot comparison)
edna.specaccum.rand <- specaccum(edna_reads, "random", permutations = 1000)
uvc.specaccum.rand <- specaccum(data_UVC, "random", permutations = 1000)


par(mfrow = c(1, 2))
#build a expected curve (randomization for boxplot comparison)
plot(uvc.specaccum.rand,ci.type="poly", col="cornflowerblue",ylab="Sobs", xlab="Samples", lwd=2, ci.lty=0, ci.col="papayawhip", main="UVC")
plot(edna.specaccum.rand,ci.type="poly", col="cornflowerblue",ylab="Sobs", xlab="Samples", lwd=2, ci.lty=0, ci.col="papayawhip", main="eDNA metabarcoding")

```


```{r}
#rarefaction OTUs richness
(S<-specnumber(edna_reads))
(raremax <- min(rowSums(edna_reads)))
(Srare<-rarefy(edna_reads,raremax))
plot(S, Srare, xlab = "Observed No. of Species", ylab = "Rarefied No. of Species")
abline(0, 1)
rarecurve(edna_reads, step = 20, xlab = "Reads",ylab = "Richness", sample = raremax, col = "cornflowerblue", cex = 0.6,main="Rarefaction eDNA")
```

```{r}
#rarefaction species richness

(S.UVC<-specnumber(data_UVC))
(raremax.UVC <- min(rowSums(data_UVC)))
(Srare.UVC<-rarefy(data_UVC,raremax.UVC))
plot(S.UVC, Srare.UVC, xlab = "Observed No. of Species", ylab = "Rarefied No. of Species")
abline(0, 1)
rarecurve(data_UVC, step = 20, xlab = "Individuals",ylab = "Richness", sample = raremax.UVC, col = "cornflowerblue", cex = 0.6, main="Rarefaction UVC")
```

```{r}
#get richness estimators (for each sample, cumulative)
pool.trasect.rare <- poolaccum(edna_reads)
#plot all: obs richness and  estimators
plot(pool.trasect.rare)
```

```{r}
summary(pool.trasect.rare)
```

```{r}
#get richness estimators (for each sample, cumulative)
pool.trasect.rare.UVC <- poolaccum(data_UVC)
#plot all: obs richness and  estimators
plot(pool.trasect.rare.UVC)

```

```{r}
summary(pool.trasect.rare.UVC)
```

----------------
JACCARD DISTANCE MATRIX
----------------
```{r}
spe.jac <- vegdist(edna_reads,method="jac", binary=TRUE)
```

```{r}
UVC.jac <- vegdist(data_UVC,method="jac", binary=TRUE)
```
----------------
BETADISPER
----------------

1. Determine regions as factors
```{r}
#2 regions
region1<-as.factor(env$region)
```

2. We now have to use the distance matrix to calculate the multivariate dispersions (variances; average distance to centroids):

```{r}
mod1 <- betadisper(spe.jac, region1)
```

```{r}
anova(mod1)
```

```{r}
adonis(spe.jac ~ region1, perm=9999)
```

```{r}
plot(mod1)
```

```{r}
mod2 <- betadisper(UVC.jac, region1)
```

As mentioned before, we have to use group dispersions to perform an ANOVA test. We will see that the ANOVA's p-value is not significant meaning that group dispersions are homogenous ("Null hypothesis of no difference in dispersion between groups"; https://www.rdocumentation.org/packages/vegan/versions/2.4-2/topics/betadisper).

Group dispersions are not homogeneous, difference in dispersion between groups

```{r}
anova(mod2)
```

```{r}
plot(mod2)
```

------------
PERMANOVA
------------
```{r}
adonis(spe.jac ~ region1, perm=9999)
```

```{r}
adonis(UVC.jac ~ region1, perm=9999)
```

---------------
nMDS
_______________


```{r}
#Data
set.seed(199)
#Ordination
NMDS_eDNA<-metaMDS(edna_reads, distance = "jac", binary=TRUE, autotransform = FALSE, k=2, trymax=5000)
NMDS_UVC<-metaMDS(data_UVC, distance = "jac", binary=TRUE, autotransform = FALSE, k=2, trymax=5000)

```



```{r}
name<-c("POR","BSS","ANI","SDI","SCR","MAT","CAT","MON","DAN","CAR","COR","PUL","ILD","SMAR","TOR","NOL","PMA", "FRA","LOR","IA-I","LOB","PAT", "TIB","EST")

#UVC
par(mfrow=c(2,1), mar=c(3, 3, 1, 1))
par(mgp=c(1.7,1,0)) 
plot(NMDS_UVC$points, col=region1,  pch = 16, 
     cex=0.7, xlab="NMDS 1", ylab="NMDS 2", cex.lab=0.7, 
     cex.axis=0.5, font.lab=1, xlim = c(-1.4,1.2),ylim = c(-1.05,0.7),family="Helvetica")
title("A) UVC", line=0.2,adj=0, font=1, cex.main=0.9, col.main="black", family="Helvetica")
text(NMDS_UVC$points+0.06, labels=name, cex=0.5, col="black", font=1)
#text(-1.2,0.2,labels="North", cex=0.9,font=4,col="grey")
#text(0.85,0.17,labels="Central", cex=0.9,font=4,col="grey")
#connect the points that belong to the same region with ordispader
ordispider(NMDS_UVC, groups = region1,lty=3, label=F,cex=0.5,font=6,col="grey1")
#ordiellipse(NMDS_UVC, groups=region2, lty=2, col="grey", label=F)
#Add legend
txt<-c("Stress 0.109", "ANOSIM R: 0.607","p: 0.0009")
legend('bottomleft', txt , cex=0.6, bty="n")

#EDNA
par(mgp=c(1.7,1,0)) 
plot(NMDS_eDNA$points, col=region1,  pch = 16, cex=0.7, xlab="NMDS 1", ylab="NMDS 2", cex.lab=0.7, cex.axis=0.5, font.lab=1,xlim = c(-0.4,0.4),ylim = c(-0.35,0.34),)
title("B) eDNA", line=0.2,adj=0, font=2, cex.main=0.9, col.main="black",family="Helvetica")
text(NMDS_eDNA$points+0.025, labels=name, cex=0.5, col="black", font=1)
#text(-0.32,0.15,labels="North", cex=0.9,font=4, col="grey")
#text(0.32,-0.18,labels="Central", cex=0.9,font=4,col="grey")
#connect the points that belong to the same region with ordispader
ordispider(NMDS_eDNA, groups = region1,lty=3, label=F,cex=0.5,font=4, col="grey1")
#ordiellipse(NMDS_eDNA, groups=region2, lty=2, col="grey", label=F)
#Add legend
txt<-c("Stress 0.217", "ANOSIM R: 0.202","p: 0.035")
legend('bottomleft', txt ,cex=0.6, bty="n")
```

----------------
ANOSIM
----------------

```{r}
fit_S_UVC<-anosim(data_UVC, region1, permutation=1000, distance = "jac")
summary(fit_S_UVC)
```

```{r}
anosim(data_UVC,grouping = region1, distance = "jac")
```

----------------------
MANTEL
----------------------

Geographic distance matrix calculator
```{r}
# Euclidean distance matrix on spatial coordinates (2D)
geo.dist.matrix <- dist(spa, method="euclidean")
```

eDNA Jaccard *geodist
```{r}
spe.jac<-vegdist(edna_reads,method = "jaccard", binary = TRUE)
(spe.vs.geo = mantel(spe.jac,geo.dist.matrix, method = "spearman", permutations = 9999, na.rm = TRUE))
```

UVC Jaccard*geodist
```{r}
UVC.jac<-vegdist(data_UVC, method="jaccard", binary = TRUE)
(UVC.jac.geodist = mantel(UVC.jac,geo.dist.matrix, method = "spearman", permutations = 9999, na.rm = TRUE))
```

-----------------------------------------------
Species richness comparisons
-----------------------------------------------

```{r}
library(ggplot2)
library(dplyr)
library(forcats)
```

```{r}
Sobs_m_r<-read.csv("Sobs_method_region.csv",header=TRUE, colClasses = c("factor","factor","factor","numeric"))
head(Sobs_m_r)
```

```{r}
t.test(Sobs_m_r$Sobs)
```

```{r}
two.way.interaction<-aov(Sobs_m_r$Sobs~Sobs_m_r$Survey.method*Sobs_m_r$GOC.region)
summary(two.way.interaction)
```

```{r}
two.way.sum<-aov(Sobs_m_r$Sobs~Sobs_m_r$Survey.method+Sobs_m_r$GOC.region)
summary(two.way.sum)
```


```{r} 
tukey.two.way<-TukeyHSD(two.way.sum)
tukey.two.way
```

```{r}
Richness_method<-read.csv("Sobs.csv",header=TRUE)

p13<-Sobs_m_r%>%
ggplot(aes(x=fct_reorder(Survey.method,Sobs, .desc=T),y=Sobs, fill=Survey.method)) +
  geom_boxplot() + 
   geom_jitter(width=0.3, alpha=1) +
   xlab("GC region")+
   theme(axis.title.x= element_blank(), axis.title = element_text(face = "bold"), legend.title=element_blank(),legend.position = ("none"))+
   facet_wrap(~GOC.region)

p13 + scale_fill_grey(start=0.8, end=0.5) + theme_linedraw()

#text=element_text(size=11, family="Helvetica"),axis.title = element_text(face = "bold"),axis.text.x=element_text(size=11),
```

```{r}

p14<-Sobs_m_r%>%
ggplot(aes(x=fct_reorder(GOC.region,Sobs, .desc=T),y=Sobs, fill=GOC.region)) +
  geom_boxplot() + 
   geom_jitter(width=0.3, alpha=1) +
   xlab("Survey method")+
   theme(axis.title.x=element_blank(),axis.title = element_text(face = "bold"),legend.title=element_blank(),legend.position = ("none"))+
   facet_wrap(~Survey.method)

p14 + scale_fill_grey(start=0.9, end=0.5) + theme_linedraw()

#axis.text.x= element_blank(),
```


